\section{Experimental Setup}
\label{sec:setup}

\paragraph{Dataset.}
We train and evaluate on HeteroMini-v1, a four-domain pretraining corpus of approximately
6.6M tokens from Web (FineWeb-Edu), Wikipedia, Code (The Stack, Python), and Literature
(Project Gutenberg), tokenized with a shared byte-level BPE vocabulary of 8,000 tokens.
Documents are kept contiguous; evaluation uses a matched held-out split with no training
overlap.

\paragraph{Model and training.}
We use SR-Core with $n = 64$ blocks, $k = 8$ active blocks per token, and $R = 6$
recursion steps (${\sim}19$M parameters, $d = 256$, block hidden dimension 512).
The dense quality baseline is a 24-layer Transformer (\textsc{dense\_d24},
${\sim}100$M parameters).
All models use AdamW with a cosine schedule and a deep-supervision LM loss.
SR-Core models are trained for 15,000 steps as a shared base, then continued for 2,000
steps with the entropy objective at $\lambda \in \{0.001, 0.003, 0.005, 0.007\}$; the
unregularized continuation (\textsc{ctrl}) provides the direct ablation baseline.
The dense baseline trains end-to-end for 17,000 steps.
Cross-seed replication uses an independently initialized base; we report seed~1 results
for $\lambda \in \{0.003, 0.005\}$.

\paragraph{Negative controls.}
\emph{Softfull} applies a soft Jaccard penalty on router probability vectors across
consecutive tokens; \emph{softsharp\_a2} adds temperature sharpening ($\alpha = 2$);
\emph{reduced\_noise} uses lower Gumbel noise ($\sigma = 0.1$).
All controls train for 2,000 steps from the shared 15k base.

\paragraph{Evaluation.}
Routing metrics are measured on 40 held-out batches via \texttt{eval\_compare.py}:
router entropy $H(\mathbf{p})$, unique core combinations, hard-overlap Jaccard, and LRU
bytes/token at cache capacities $K \in \{8,16,24,32\}$.
Per-domain held-out loss is measured via paired runs with the dense baseline as a
consistent anchor.
Offloading simulation uses an LRU cache trace over 1,024 held-out sequences.
