\section{Related Work}
\label{sec:related}

\paragraph{Mixture of Experts.}
Sparse gating routes tokens to expert feed-forward networks
\citep{shazeer2017moe,fedus2022switch}.
Standard MoE architectures activate a different expert subset at each layer, accumulating
a transfer budget proportional to depth; SR-Core replaces per-layer routing with a single
shared bank and a fixed working set $\WS = k$, independent of recursion depth.
Load-balancing losses \citep{lepikhin2021gshard,fedus2022switch} address expert collapse;
the entropy objective addresses over-diffuse routing --- a distinct failure mode.

\paragraph{Looped and recurrent depth.}
The Universal Transformer \citep{dehghani2019universal} and Deep Equilibrium Models
\citep{bai2019deq} apply shared parameters repeatedly.
SR-Core fixes routing at $r = 1$, giving a predictable active-block set at inference time
rather than a variable computation graph --- a property these models do not address.

\paragraph{Parameter offloading.}
FlexGen \citep{sheng2023flexgen} and DeepSpeed Inference \citep{aminabadi2022deepspeed}
optimize layer-offloading schedules for dense transformers; SR-Core reduces \emph{what}
is loaded per token.
The two approaches are complementary.

\paragraph{Routing regularization.}
MoE auxiliary objectives include load-balancing losses \citep{fedus2022switch}, z-loss
\citep{zoph2022stmoe}, and router noise \citep{shazeer2017moe}.
These target dead experts or training instability; none control the cross-token cache
locality that the entropy objective addresses.

\paragraph{Weight sharing.}
SHA-RNN \citep{merity2019sharnn} and related architectures reuse the same weights
structurally; SR-Core adds content-dependent routing, keeping the active count fixed at
$k$ while varying \emph{which} weights are active per token.
